Comparing Different Planned Missingness Designs in Longitudinal Studies

### 2018-05-15 Documentation/manuscript.Rnw
Manuscript is now being writted in KNITR.

\begin{table}[h!]
	\centering
	\caption{The Three Form PM Design}
	\label{tab:table10}
	\setlength{\tabcolsep}{1cm}
	\begin{tabular}{c|cccc}
		\toprule
		& \multicolumn{4}{c}{Block} \\
		Form & X & A & B & C \\
		\midrule
		1 & 1 & 1 & 1 & 0 \\
		2 & 1 & 1 & 0 & 1 \\
		3 & 1 & 0 & 1 & 1 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

The study employed an observational cohort design, following sixty participants over forty-five days and sending them a daily questionnaire. Questions in the survey concerned livelihood insecurities from the previous day as well as alcohol use. The analysis used eight particular questions, two binary questions on housing insecurity, four binary questions on food insecurity, and two questions concerning alcohol use  (one binary variable and one count variable). The exact questions can be found in Table \ref{tab:table1}. \par
\begin{table}[t]
	\centering
	\caption{Survey Questions}
	\label{tab:table1}
	\begin{tabular}{cp{6.5cm}c}
		\toprule
		Item Number & Question & Responses\\
		\midrule
		1 & I worried about having a place to stay yesterday. & 1.Yes $\>$ 2.No \\
		2 & I could not get to where I needed to go yesterday. & 1.Yes $\>$ 2.No \\
		3 & I worried about my food running out yesterday. & 1.Yes $\>$ 2.No \\
		4 & I ate less than I needed to yesterday. & 1.Yes $\>$ 2.No \\
		5 & I was hungry, but did not eat because I couldn't afford food. & 1.Yes $\>$ 2.No \\
		6 & I got food from a pantry, church, agency, friend or the street yesterday. & 1.Yes $\>$ 2.No \\
		7 & Did you drink alcohol yesterday? & 1.Yes $\>$ 2.No \\
		8 & How many alcohol drinks did you have yesterday? & 0-20 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

>>>>

We need to talk about the other type of split design that we used here which was that X of the 8 questions are randomly omitted on each day.

% \begin{table}[h!]
% 	\centering
% 	\caption{The Three Form PM Design}
% 	\label{tab:table10}
% 	\setlength{\tabcolsep}{1cm}
% 	\begin{tabular}{c|cccc}
% 		\toprule
% 		& \multicolumn{4}{c}{Block} \\
% 		Form & X & A & B & C \\
% 		\midrule
% 		1 & 1 & 1 & 1 & 0 \\
% 		2 & 1 & 1 & 0 & 1 \\
% 		3 & 1 & 0 & 1 & 1 \\
% 		\bottomrule
% 	\end{tabular}
% \end{table}

<<echo = FALSE>>=
# Creating dataframe to visualize - one subject's questions for 5 days
split.form <- function(set, nrem = 2, bound = 0){

set[set$Day >= bound, paste0("Q",1:8)] <- t(apply(set[set$Day >= bound,paste0("Q",1:8)], 1, function(x){
    ind <- sample(1:8, nrem)
    x[ind] <- NA
    as.numeric(x)
  }))

  return(set)

}
@

<<echo = FALSE>>=
wave.des <- function(set, nmiss, ...) { # nmiss is the number of days each person would miss? 11?? Like, miss a fourth of the days?
  id <- sample(2:45, size=nmiss)
  for (i in 1:nmiss) {
    set[set$Day %in% id[i], c('Q1','Q2','Q3','Q4','Q5','Q6','Q7','Q8')] <- NA
  }
  return(set)
}
@

<<warning=FALSE, echo = FALSE>>=
cols <- c('Q1','Q2','Q3','Q4','Q5','Q6','Q7','Q8')
fill <- matrix(1, 45, 8)
dat <- data.frame(fill)
colnames(dat) <- cols
dat$Day <- seq.int(nrow(dat))
set.seed(1234)
split.dat <- split.form(dat)[1:8]
#split.dat
@

<<warning=FALSE,message=FALSE,echo=FALSE>>=
library(ggplot2)
library(dplyr)
library(reshape2)
datmelt <- split.dat[1:8] %>% is.na %>% melt
colnames(datmelt) <- c("Day","Question","Missing")
#datmelt$block <- ifelse((datmelt$Question=="Q1")|(datmelt$Question=="Q8"),"X",ifelse((datmelt$Question=="Q2")|(datmelt$Question=="Q3"),"A",ifelse((datmelt$Question=="Q4")|(datmelt$Question=="Q5"),"B",ifelse((datmelt$Question=="Q6")|(datmelt$Question=="Q7"),"C",0))))
datmelt$color <- ifelse(datmelt$Missing == "TRUE", "Missing", "Observed")

  #ifelse(datmelt$Missing == "TRUE", "Missing", ifelse(datmelt$block == "X","X",ifelse(datmelt$block == "A","A",ifelse(datmelt$block == "B","B",ifelse(datmelt$block == "C","C","Missing")))))
@

<<warning=FALSE,echo=FALSE>>=
cols <- c("Missing" = "red", "Observed" = "skyblue3")
splitplot <- ggplot(datmelt, aes(Question, Day, fill=color)) + geom_raster() + scale_fill_manual(values = cols)  + theme(axis.text.x = element_text(angle=45, vjust=0.5),panel.background = element_rect(fill = NA)) + labs(x = "Questions Asked",y = "Day",fill="Block") + scale_y_reverse() + theme(legend.title=element_blank(),legend.position="none")

#+ geom_vline(xintercept=seq(1.5, 7.5, 2), lwd=1, colour="white") + geom_hline(yintercept=seq(1.5, 4.5, 1), lwd=1, colour="white")
@

<<warning=FALSE,echo=FALSE>>=
cols <- c('Q1','Q2','Q3','Q4','Q5','Q6','Q7','Q8')
fill <- matrix(1, 45, 8)
dat <- data.frame(fill)
colnames(dat) <- cols
dat$Day <- seq.int(nrow(dat))
set.seed(1234)
alternate.split.dat <- split.form(dat, bound = 3)[1:8]
#alternate.split.dat
@

<<warning=FALSE,message=FALSE,echo=FALSE>>=
library(ggplot2)
library(dplyr)
library(reshape2)
datmelt <- alternate.split.dat[1:8] %>% is.na %>% melt
colnames(datmelt) <- c("Day","Question","Missing")
#datmelt$block <- ifelse((datmelt$Question=="Q1")|(datmelt$Question=="Q8"),"X",ifelse((datmelt$Question=="Q2")|(datmelt$Question=="Q3"),"A",ifelse((datmelt$Question=="Q4")|(datmelt$Question=="Q5"),"B",ifelse((datmelt$Question=="Q6")|(datmelt$Question=="Q7"),"C",0))))
datmelt$color <- ifelse(datmelt$Missing == "TRUE", "Missing", "Observed")

  #ifelse(datmelt$Missing == "TRUE", "Missing", ifelse(datmelt$block == "X","X",ifelse(datmelt$block == "A","A",ifelse(datmelt$block == "B","B",ifelse(datmelt$block == "C","C","Missing")))))
@

<<warning=FALSE,echo=FALSE>>=
cols <- c("Missing" = "red", "Observed" = "skyblue3")
alternatesplitplot <- ggplot(datmelt, aes(Question, Day, fill=color)) + geom_raster() + scale_fill_manual(values = cols)  + theme(axis.text.x = element_text(angle=45, vjust=0.5),panel.background = element_rect(fill = NA)) + labs(x = "Questions Asked",y = "Day",fill="Block") + scale_y_reverse() + theme(legend.title=element_blank(),legend.position="none")

#+ geom_vline(xintercept=seq(1.5, 7.5, 2), lwd=1, colour="white") + geom_hline(yintercept=seq(1.5, 4.5, 1), lwd=1, colour="white")
@

<<echo=FALSE>>=
set.seed(1234)
wave.dat <- wave.des(dat,11)
#wave.dat
@

<<echo=FALSE>>=
library(ggplot2)
library(dplyr)
library(reshape2)
datmelt <- wave.dat[1:8] %>% is.na %>% melt
colnames(datmelt) <- c("Day","Question","Missing")
#datmelt$block <- ifelse((datmelt$Question=="Q1")|(datmelt$Question=="Q8"),"X",ifelse((datmelt$Question=="Q2")|(datmelt$Question=="Q3"),"A",ifelse((datmelt$Question=="Q4")|(datmelt$Question=="Q5"),"B",ifelse((datmelt$Question=="Q6")|(datmelt$Question=="Q7"),"C",0))))
datmelt$color <- ifelse(datmelt$Missing == "TRUE", "Missing", "Observed")

  #ifelse(datmelt$Missing == "TRUE", "Missing", ifelse(datmelt$block == "X","X",ifelse(datmelt$block == "A","A",ifelse(datmelt$block == "B","B",ifelse(datmelt$block == "C","C","Missing")))))
@

<<echo=FALSE>>=
cols <- c("Missing" = "red", "Observed" = "skyblue3")
waveplot <-  ggplot(datmelt, aes(Question, Day, fill=color)) + geom_raster() + scale_fill_manual(values = cols)  + theme(axis.text.x = element_text(angle=45, vjust=0.5),panel.background = element_rect(fill = NA)) + labs(x = "Questions Asked",y = "Day",fill="Block") + scale_y_reverse() + theme(legend.title=element_blank(),legend.position="none")
@

<<misslow, fig.cap = "Low missingness", echo = FALSE, message = FALSE>>=
require(gridExtra)
grid.arrange(splitplot, alternatesplitplot, waveplot, ncol=3)
@

Figure \ref{fig:misslow} is

@@

The study employed an observational cohort design, following sixty participants over forty-five days and sending them a daily questionnaire. Questions in the survey concerned livelihood insecurities from the previous day as well as alcohol use. The analysis used eight particular questions, two binary questions on housing insecurity, four binary questions on food insecurity, and two questions concerning alcohol use  (one binary variable and one count variable). The exact questions can be found in Table \ref{tab:table1}. \par

\begin{table}[t]
	\centering
	\caption{Survey Questions}
	\label{tab:table1}
	\begin{tabular}{cp{6.5cm}c}
		\toprule
		Item Number & Question & Responses\\
		\midrule
		1 & I worried about having a place to stay yesterday. & 1.Yes $\>$ 2.No \\
		2 & I could not get to where I needed to go yesterday. & 1.Yes $\>$ 2.No \\
		3 & I worried about my food running out yesterday. & 1.Yes $\>$ 2.No \\
		4 & I ate less than I needed to yesterday. & 1.Yes $\>$ 2.No \\
		5 & I was hungry, but did not eat because I couldn't afford food. & 1.Yes $\>$ 2.No \\
		6 & I got food from a pantry, church, agency, friend or the street yesterday. & 1.Yes $\>$ 2.No \\
		7 & Did you drink alcohol yesterday? & 1.Yes $\>$ 2.No \\
		8 & How many alcohol drinks did you have yesterday? & 0-20 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

<<>>=
load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_high.RData")
high <- d

load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_med.RData")
med <- d

load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_low.RData")
low <- d

library(xtable)
print(xtable(cbind("","","",rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])),digits= 3),include.rownames=FALSE)

load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_high_hightime1.RData")
high <- d

load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_med_hightime1.RData")
med <- d

load("~/Dropbox/designedMissingnessGit/Results/tables20180509_60_1_low_hightime1.RData")
low <- d

library(xtable)
print(xtable(cbind("","","",rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])),digits= 3),include.rownames=FALSE)

@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon May 14 21:02:34 2018
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrr}
  \hline
Design & \% Missingness  & Parameter & Coverage & Bias & MSE & FMI \\ 
  \hline
 &  &  & 0.989 & 0.042 & 0.018 & 0.002 \\ 
   &  &  & 0.984 & -0.125 & 0.731 & 0.270 \\ 
   &  &  & 0.934 & 0.008 & 0.024 & 0.000 \\ 
   &  &  & 0.943 & -0.000 & 0.000 & 0.001 \\ 
   \hline
   &  &  & 0.986 & 0.040 & 0.019 & 0.012 \\ 
   &  &  & 0.993 & -0.168 & 0.750 & 0.557 \\ 
   &  &  & 0.932 & 0.008 & 0.024 & 0.002 \\ 
   &  &  & 0.940 & -0.000 & 0.000 & 0.008 \\ 
   \hline
   &  &  & 0.984 & 0.036 & 0.027 & 0.030 \\ 
   &  &  & 0.991 & -0.713 & 9.933 & 0.625 \\ 
   &  &  & 0.932 & 0.009 & 0.027 & 0.005 \\ 
   &  &  & 0.939 & -0.000 & 0.000 & 0.021 \\ 
   \hline
   &  &  & 0.988 & 0.042 & 0.018 & 0.002 \\ 
   &  &  & 0.977 & -0.114 & 0.696 & 0.250 \\ 
   &  &  & 0.933 & 0.008 & 0.024 & 0.000 \\ 
   &  &  & 0.943 & -0.000 & 0.000 & 0.001 \\ 
   \hline
   &  &  & 0.988 & 0.040 & 0.018 & 0.010 \\ 
   &  &  & 0.994 & -0.167 & 0.782 & 0.549 \\ 
   &  &  & 0.931 & 0.009 & 0.024 & 0.002 \\ 
   &  &  & 0.944 & -0.000 & 0.000 & 0.006 \\ 
   \hline
   &  &  & 0.983 & 0.036 & 0.022 & 0.027 \\ 
   &  &  & 0.990 & -0.686 & 10.509 & 0.665 \\ 
   &  &  & 0.928 & 0.009 & 0.026 & 0.007 \\ 
   &  &  & 0.946 & -0.000 & 0.000 & 0.023 \\ 
   \hline
   &  &  & 0.991 & 0.040 & 0.018 & 0.006 \\ 
   &  &  & 0.998 & -0.116 & 0.545 & 0.458 \\ 
   &  &  & 0.932 & 0.008 & 0.024 & 0.001 \\ 
   &  &  & 0.941 & -0.000 & 0.000 & 0.003 \\ 
   \hline
   &  &  & 0.991 & 0.029 & 0.018 & 0.055 \\ 
   &  &  & 0.988 & -0.112 & 0.345 & 0.752 \\ 
   &  &  & 0.934 & 0.009 & 0.024 & 0.007 \\ 
   &  &  & 0.951 & -0.000 & 0.000 & 0.032 \\ 
   \hline
   &  &  & 0.994 & -0.003 & 0.022 & 0.258 \\ 
   &  &  & 0.988 & -0.105 & 0.284 & 0.903 \\ 
   &  &  & 0.937 & 0.011 & 0.025 & 0.039 \\ 
   &  &  & 0.960 & -0.000 & 0.000 & 0.175 \\ 
   \hline
\end{tabular}
\end{table}
